---
title: "Assignment_Week2"
author: "Jan Willem"
date: "22-2-2021"
output: html_document
---

## Assignment

1 Code for reading in the dataset and/or processing the data  
2 Histogram of the total number of steps taken each day  
3 Mean and median number of steps taken each day  
4 Time series plot of the average number of steps taken  
5 The 5-minute interval that, on average, contains the maximum number of steps  
6 Code to describe and show a strategy for imputing missing data  
7 Histogram of the total number of steps taken each day after missing values are imputed  
8 Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends  
9 All of the R code needed to reproduce the results (numbers, plots, etc.) in the report  


### 1 Code for reading in the dataset and/or processing the data

Download the data

```{r}
options(scipen = 999) 
#Download and unzip files
download.file(url = "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
              , destfile = paste(getwd(), "repdata_data_activity.zip", sep = "/"))
unzip(zipfile = "repdata_data_activity.zip")

```
Check for pacakges that are needed 
```{r echo=FALSE, message=FALSE, warning=FALSE}
## check and if necessary install packages
packages <- c("dplyr", "ggplot2", "gridExtra", "imputeTS", "lubridate")
install.packages(setdiff(packages, rownames(installed.packages())))
##loading packages  
lapply(packages, library, character.only = TRUE)
```


Read file and look at the head of the file
```{r}
activity<-read.csv(file="activity.csv" , header=TRUE, sep=",")
head(activity)
```

Change datatype dat to date()
```{r}
activity$date<-as.Date(activity$date)
str(activity)
```
### 2 Histogram of the total number of steps taken each day

```{r warning=FALSE}
steps_each_day<-activity%>%
                group_by(date)%>%
                summarise_at(vars(steps),list(sum=sum))



ggplot(data=steps_each_day, aes(y=sum, x=date)) + 
      geom_col()+ 
      labs(title="Total number of steps taken each day") +
      labs(x="Date", y="Total steps")

```

### 3 Mean and median number of steps taken each day

```{r}
mediaan<-median(steps_each_day$sum, na.rm=TRUE)
gemiddeld<-round(mean(steps_each_day$sum, na.rm=TRUE),2)

mediaan
gemiddeld
```


The median of the number of steps taken each day is:  `r mediaan`.
The mean of the number of steps is: `r gemiddeld`

### 4 Time series plot of the average number of steps taken


```{r warning=FALSE}
steps_each_day_mean<-steps_each_day%>%
                    group_by(date)%>%
                    summarise_at(vars(sum),list(mean=mean))

ggplot(steps_each_day_mean, aes(date, mean)) + geom_line() 

```

### 5 The 5-minute interval that, on average, contains the maximum number of steps

```{r}
avg_most_steps<-activity%>%                            # data frame
                group_by(interval) %>%                 # group indicator
                summarise_at(vars(steps),              # column
                             list(gemiddeld = mean), na.rm=TRUE)


subset(avg_most_steps,gemiddeld==max(avg_most_steps$gemiddeld))

```

### 6 Code to describe and show a strategy for imputing missing data


First of all w heave to check if and where missing data is.
So we check all variables if any missing data exists
```{r}
s<-anyNA(activity$steps)
d<-anyNA(activity$date)
i<-anyNA(activity$interval)

```
Are there NA's in activity$step `r s`    

Are there NA's in activity$date `r d`  

Are there NA's in activity$interval `r i`  

Because this is a timeseries, I choose for the imputeTS package

First find out where the missing data is:
```{r}
#where are the missing data
ggplot_na_distribution(activity$steps)
```

Ok now its time to see what different impute methods are doing
```{r}
imp_mean <- na_mean(activity$steps)
p1<-ggplot_na_imputations(activity$steps, imp_mean, subtitle= "Visualization of missing value replacements\nmethod=mean")

imp_interpolation <- na_interpolation(activity$steps)
p2<-ggplot_na_imputations(activity$steps, imp_interpolation, subtitle= "Visualization of missing value replacements\nmethod=interpolation")

imp_kalman <- na_kalman(activity$steps)
p3<-ggplot_na_imputations(activity$steps, imp_kalman, subtitle= "Visualization of missing value replacements\nmethod=kalman")

imp_ma <- na_ma(activity$steps)
p4<-ggplot_na_imputations(activity$steps, imp_ma, subtitle= "Visualization of missing value replacements \nmethod=ma")


#Overview of plots
grid.arrange(p1, p2, p3, p4, nrow = 2)

```


It looks like the method mean is the best method for imputing, so that will it be.
```{r}
imp_mean <- na_mean(activity$steps)
ggplot_na_imputations(activity$steps, imp_mean, subtitle= "Visualization of missing value replacements, method=mean")
```

### 7 Histogram of the total number of steps taken each day after missing values are imputed

```{r}
activity_imp<-cbind(activity, imp_mean)
steps_each_day_imp<-activity_imp%>%
                group_by(date)%>%
                summarise_at(vars(imp_mean),list(sum=sum))

ggplot(data=steps_each_day_imp, aes(y=sum, x=date)) + 
      geom_col()+ 
      labs(title="Total number of steps taken each day IMPUTED") +
      labs(x="Date", y="Total steps")

```


### 8 Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}
avg_most_steps_comp<-activity%>%   
                     mutate(weekday= ifelse((wday(date)==1 | wday(date)==7), "weekends", "weekdays")) %>%
                     group_by(weekday,interval) %>%                 
                     summarise_at(vars(steps), list(gemiddeld = mean), na.rm=TRUE)


d <- ggplot(avg_most_steps_comp, aes(interval,gemiddeld)) + geom_line() 
d + facet_wrap(~ weekday)
````


### 9 All of the R code needed to reproduce the results (numbers, plots, etc.) in the report


Ha,  see above
